---
title: "Landscape Integrity Index (LII) - Model Validation"
author: "Liling Lee"
date: "11/02/2019"
output:
  prettydoc::html_pretty:
    theme: architect
    highlight: github
  html_notebook: default

fontsize: 14pt
---

## Introduction
The model was validated by (1) comparing the LII values with the values from the Landscape Condition Map (LCM) developed by Hak and Comer (2017) using a linear regression model, and (2) comparing the LII values in protected areas to the LII values in multiple-use areas using the Protected Areas Database of the United States (PAD-US) data from U.S. Geology Survey (USGS) Gap Analysis Project (GAP) and Welch's two sample t-test. The linear regression would show if there was a relationship between Landscape Integrity Index and Landscape Condition Map. And the Welch's two sample t-test would indicate if the mean LII value would be different between protected areas and multiple-use Areas. 

```{r}
# Set working directory
wd <- setwd("C:/LII")

# Install and load readxl, ggplot2, dplyr, tidyr, magrittr, gridExtra, e1071, kableExtra, and knitr packages

if (!requireNamespace("readxl", quietly = TRUE))
  install.packages("readxl")
if (!requireNamespace("ggplot2", quietly = TRUE))
  install.packages("ggplot2")
if (!requireNamespace("dplyr", quietly = TRUE))
  install.packages("dplyr")
if (!requireNamespace("tidyr", quietly = TRUE))
  install.packages("tidyr")
if (!requireNamespace("magrittr", quietly = TRUE))
  install.packages("magrittr")
if (!requireNamespace("gridExtra", quietly = TRUE))
  install.packages("gridExtra")
if (!requireNamespace("e1071", quietly = TRUE))
  install.packages("e1071")
if (!requireNamespace("kableExtra", quietly = TRUE))
  install.packages("kableExtra")
if (!requireNamespace("knitr", quietly = TRUE))
  install.packages("knitr")

library(readxl)
library(ggplot2)      # plotting & data
library(dplyr)        # data manipulation
library(tidyr)        # data re-shaping
library(magrittr)     # pipe operator
library(gridExtra)    # provides side-by-side plotting
library(e1071)        # for skewness function
library(kableExtra)
library(knitr)

```

```{r}
# Import data
LII_LCM <- read_excel("C:/LII/ModelValidation.xlsx", sheet = "LII_LCM")
LII_PADUS_100 <- read_excel("C:/LII/ModelValidation.xlsx", sheet = "LII_PADUS_100")

```

```{r}
# Plot scatter plot to visualize the relationship
scatter.smooth(x=LII_LCM$LCM_Value, y=LII_LCM$LII_Value, main="LCM ~ LII")

```

```{r}
# Plot box plot to check for outliers
par(mfrow=c(1, 2))  # divide graph area in 2 columns

boxplot(LII_LCM$LCM_Value, main="LCM", sub=paste("Outlier rows: ", boxplot.stats(LII_LCM$LCM_Value)$out))  # box plot for 'LCM'

boxplot(LII_LCM$LII_Value, main="LII", sub=paste("Outlier rows: ", boxplot.stats(LII_LCM$LII_Value)$out))  # box plot for 'LII'
```

```{r}
# Plot density plot to check if response variable is close to normal
par(mfrow=c(1, 2))  # divide graph area in 2 columns

plot(density(LII_LCM$LCM_Value), main="LCM", ylab="Frequency", sub=paste("Skewness:", round(e1071::skewness(cars$speed), 2)))  # density plot for 'LCM'

polygon(density(LII_LCM$LCM_Value), col="red")

plot(density(LII_LCM$LII_Value), main="LII", ylab="Frequency", sub=paste("Skewness:", round(e1071::skewness(cars$dist), 2)))  # density plot for 'LII'

polygon(density(LII_LCM$LII_Value), col="red")

```

```{r}
# Calculate correlation between LCM value and LII value
cor(LII_LCM$LCM_Value, LII_LCM$LII_Value)
```

```{r}
# Build the linear regression model
LII_LCM_lm <- lm(LII_LCM$LII_Value ~ LII_LCM$LCM_Value, data=LII_LCM)  # build linear regression model on full data
print(LII_LCM_lm)
summary(LII_LCM_lm)

```

```{r}
str(summary(LII_LCM_lm))
```

```{r}
summary(LII_LCM_lm)$r.squared
```


```{r}
# Plot linear regression line
plot(LII_LCM$LCM_Value, LII_LCM$LII_Value, pch = 16, cex = 1.3, col = "blue", main = "Linear Regression Line of LII and LCM Values", xlab = "LCM Value", ylab = "LII Value")

abline(LII_LCM_lm)
```

```{r}
# Plot linear regression line
plot(LII_LCM$LCM_Value, LII_LCM$LII_Value, pch = 16, cex = 1.3, col = "blue", main = "", xlab = "", ylab = "")

abline(LII_LCM_lm)
```

```{r}
# Calculate the AIC and BIC for model goodness of fit
AIC(LII_LCM_linearMod) 
BIC(LII_LCM_linearMod) 

```

```{r}
LII_PADUS_100_df <- LII_PADUS_100 %>%
  filter(group == "Protected" | group == "Multiple Use") %>%
  select(group, RASTERVALU)

summary(LII_PADUS_100_df %>% filter(group == "Protected") %>% .$RASTERVALU)

summary(LII_PADUS_100_df %>% filter(group == "Multiple Use") %>% .$RASTERVALU)

```

```{r}
ggplot(LII_PADUS_100_df, aes(group, RASTERVALU)) + geom_boxplot()

```

```{r}
LII_PADUS_ttest <- t.test(RASTERVALU ~ group, data = LII_PADUS_100_df)
LII_PADUS_ttest

```

```{r}
summary(LII_PADUS_ttest)

```