---
title: "Landscape Integrity Index (LII) - Landscape Metrics"
author: "Liling Lee"
date: "09/18/2019"
output:
  prettydoc::html_pretty:
    theme: architect
    highlight: github
  html_notebook: default

fontsize: 14pt
---

## Introduction

The purpose of the model is to create the landscape metrics variables for the Landscape Integrity Index (LII) using the 'landscapemetrics' package and NLCD data. 

## Software Requirements 

* R v.3.5.3

## Data Preparation and Pre-Processing

**Process:**

1. Install and load the necessary packages and set the working directory.

**Note:** User needs to change the working directory.
s
```{r}
# Set working directory
wd <- setwd("C:/LII")

# Install and load raster, rgdal, sp, kableExtra, knitr, and landscapemetrics packages

if (!requireNamespace("raster", quietly = TRUE))
  install.packages("raster")
if (!requireNamespace("rgdal", quietly = TRUE))
  install.packages("rgdal")
if (!requireNamespace("sp", quietly = TRUE))
  install.packages("sp")
if (!requireNamespace("kableExtra", quietly = TRUE))
  install.packages("kableExtra")
if (!requireNamespace("knitr", quietly = TRUE))
  install.packages("knitr")
if (!requireNamespace("landscapemetrics", quietly = TRUE))
  install.packages("landscapemetrics")

library(raster)
library(rgdal)
library(sp)
library(kableExtra)
library(knitr)
library(landscapemetrics)

```

**Data Needed:**
* Land Cover Data: NLCD
    + **Note:** The resolution of a raster cell has to be in meters, since the package converts units internally and returns results in either meters, square meters or hectares. 


**Projection:** 
NAD 1983 UTM Zone 13N

**Note:** The user needs to change the projection that is suitable for the study area.

```{r}
# Import raster
NLCD2001 <- raster("C:/LII/landscapemetrics/NLCD2001.tif")
NLCD2004 <- raster("C:/LII/landscapemetrics/NLCD2004.tif")
NLCD2006 <- raster("C:/LII/landscapemetrics/NLCD2006.tif")
NLCD2008 <- raster("C:/LII/landscapemetrics/NLCD2008.tif")
NLCD2011 <- raster("C:/LII/landscapemetrics/NLCD2011.tif")
NLCD2013 <- raster("C:/LII/landscapemetrics/NLCD2013.tif")
NLCD2016 <- raster("C:/LII/landscapemetrics/NLCD2016.tif")

# Check landscape
check_landscape(NLCD2001)
check_landscape(NLCD2004)
check_landscape(NLCD2006)
check_landscape(NLCD2008)
check_landscape(NLCD2011)
check_landscape(NLCD2013)
check_landscape(NLCD2016)

```

## Shape Metrics: Perimeter-Area Fractal Dimension (PAFRAC)

```{r}
# Shape Metrics: Perimeter-Area Fractal Dimension (PAFRAC)
lsm_c_pafrac(NLCD2001, directions = 8, verbose = TRUE)
lsm_c_pafrac(NLCD2004, directions = 8, verbose = TRUE)
lsm_c_pafrac(NLCD2006, directions = 8, verbose = TRUE)
lsm_c_pafrac(NLCD2008, directions = 8, verbose = TRUE)
lsm_c_pafrac(NLCD2011, directions = 8, verbose = TRUE)
lsm_c_pafrac(NLCD2013, directions = 8, verbose = TRUE)
lsm_c_pafrac(NLCD2016, directions = 8, verbose = TRUE)

```

## Core Area Metrics: Coefficient of Variation of Core Area Index (CAI_CV)

```{r}
# Core Area Metrics: Coefficient of Variation of Core Area Index (CAI_CV)
lsm_c_cai_cv(NLCD2001, directions = 8, consider_boundary = FALSE, edge_depth = 1)
lsm_c_cai_cv(NLCD2004, directions = 8, consider_boundary = FALSE, edge_depth = 1)
lsm_c_cai_cv(NLCD2006, directions = 8, consider_boundary = FALSE, edge_depth = 1)
lsm_c_cai_cv(NLCD2008, directions = 8, consider_boundary = FALSE, edge_depth = 1)
lsm_c_cai_cv(NLCD2011, directions = 8, consider_boundary = FALSE, edge_depth = 1)
lsm_c_cai_cv(NLCD2013, directions = 8, consider_boundary = FALSE, edge_depth = 1)
lsm_c_cai_cv(NLCD2016, directions = 8, consider_boundary = FALSE, edge_depth = 1)

```

## Core Area Metrics: Standard Deviation of Core Area Index (CAI_SD)

```{r}
# Core Area Metrics: Standard Deviation of Core Area Index (CAI_SD)
lsm_c_cai_sd(NLCD2001, directions = 8, consider_boundary = FALSE, edge_depth = 1)
lsm_c_cai_sd(NLCD2004, directions = 8, consider_boundary = FALSE, edge_depth = 1)
lsm_c_cai_sd(NLCD2006, directions = 8, consider_boundary = FALSE, edge_depth = 1)
lsm_c_cai_sd(NLCD2008, directions = 8, consider_boundary = FALSE, edge_depth = 1)
lsm_c_cai_sd(NLCD2011, directions = 8, consider_boundary = FALSE, edge_depth = 1)
lsm_c_cai_sd(NLCD2013, directions = 8, consider_boundary = FALSE, edge_depth = 1)
lsm_c_cai_sd(NLCD2016, directions = 8, consider_boundary = FALSE, edge_depth = 1)

```

## Core Area Metrics: Coefficient of Variation of Core Area (CORE_CV)

```{r}
# Core Area Metrics: Coefficient of Variation of Core Area (CORE_CV)
lsm_c_core_cv(NLCD2001, directions = 8, consider_boundary = FALSE, edge_depth = 1)
lsm_c_core_cv(NLCD2004, directions = 8, consider_boundary = FALSE, edge_depth = 1)
lsm_c_core_cv(NLCD2006, directions = 8, consider_boundary = FALSE, edge_depth = 1)
lsm_c_core_cv(NLCD2008, directions = 8, consider_boundary = FALSE, edge_depth = 1)
lsm_c_core_cv(NLCD2011, directions = 8, consider_boundary = FALSE, edge_depth = 1)
lsm_c_core_cv(NLCD2013, directions = 8, consider_boundary = FALSE, edge_depth = 1)
lsm_c_core_cv(NLCD2016, directions = 8, consider_boundary = FALSE, edge_depth = 1)

```

## Contagion and Interspersion Metrics: Clumpy Index (CLUMPY)

```{r}
# Core Area Metrics: Clumpy Index (CLUMPY)
lsm_c_clumpy(NLCD2001)
lsm_c_clumpy(NLCD2004)
lsm_c_clumpy(NLCD2006)
lsm_c_clumpy(NLCD2008)
lsm_c_clumpy(NLCD2011)
lsm_c_clumpy(NLCD2013)
lsm_c_clumpy(NLCD2016)

```

## Diversity Metrics: Shannon's Diversity Index (SHDI)

```{r}
# Diversity Metrics: Shannon's Diversity Index (SHDI)
lsm_l_shdi(NLCD2001)
lsm_l_shdi(NLCD2004)
lsm_l_shdi(NLCD2006)
lsm_l_shdi(NLCD2008)
lsm_l_shdi(NLCD2011)
lsm_l_shdi(NLCD2013)
lsm_l_shdi(NLCD2016)

```

## Diversity Metrics: Simpson's Diversity Index (SIDI)

```{r}
# Diversity Metrics: Simpson's Diversity Index (SIDI)
lsm_l_sidi(NLCD2001, directions = 8)
lsm_l_sidi(NLCD2004, directions = 8)
lsm_l_sidi(NLCD2006, directions = 8)
lsm_l_sidi(NLCD2008, directions = 8)
lsm_l_sidi(NLCD2011, directions = 8)
lsm_l_sidi(NLCD2013, directions = 8)
lsm_l_sidi(NLCD2016, directions = 8)

```

## Diversity Metrics: Patch Richness (PR)

```{r}
# Diversity Metrics: Patch Richness (PR)
lsm_l_pr(NLCD2001)
lsm_l_pr(NLCD2004)
lsm_l_pr(NLCD2006)
lsm_l_pr(NLCD2008)
lsm_l_pr(NLCD2011)
lsm_l_pr(NLCD2013)
lsm_l_pr(NLCD2016)

```

## Diversity Metrics: Patch Richness Density (PRD)

```{r}
# Diversity Metrics: Patch Richness Density (PRD)
lsm_l_prd(NLCD2001, directions = 8)
lsm_l_prd(NLCD2004, directions = 8)
lsm_l_prd(NLCD2006, directions = 8)
lsm_l_prd(NLCD2008, directions = 8)
lsm_l_prd(NLCD2011, directions = 8)
lsm_l_prd(NLCD2013, directions = 8)
lsm_l_prd(NLCD2016, directions = 8)

```

## Diversity Metrics: Shannon's evenness index (SHEI)

```{r}
# Diversity Metrics: Shannon's evenness index (SHEI)
lsm_l_shei(NLCD2001)
lsm_l_shei(NLCD2004)
lsm_l_shei(NLCD2006)
lsm_l_shei(NLCD2008)
lsm_l_shei(NLCD2011)
lsm_l_shei(NLCD2013)
lsm_l_shei(NLCD2016)

```

## Diversity Metrics: Simpson's evenness index (SIEI)

```{r}
# Diversity Metrics: Simpson's evenness index (SIEI)
lsm_l_siei(NLCD2001, directions = 8)
lsm_l_siei(NLCD2004, directions = 8)
lsm_l_siei(NLCD2006, directions = 8)
lsm_l_siei(NLCD2008, directions = 8)
lsm_l_siei(NLCD2011, directions = 8)
lsm_l_siei(NLCD2013, directions = 8)
lsm_l_siei(NLCD2016, directions = 8)

```